### 2014-03-10 Studies/Chicago_2012/Report/Chicago_Fire.tex
Chicago Fire: minor update to report

\chapter{Introduction}
Part of the mission of the Fire Research Division at the National Institute of Standards and Technology (NIST) is to develop and apply technology, measurements and standards, improve the understanding of the behavior, prevention and control of fires to enhance fire fighting operations and equipment, fire suppression, fire investigations, and disaster response. NIST has previously used the Fire Dynamics Simulator (FDS) to provide insight on the fire development and thermal conditions of other fatality fires \cite{Madrzykowski:1,Iowa,Texas,Cook_County,Grosshandler:Station,Bryner:Charleston,barowy:texas}. The overriding objective of all of these studies is to improve the safety of fire fighters.

@@

\chapter{Fire Incident Summary}
The following account of events is based on information provided by NIST physical measurements on-scene, NIST photographs of the scene, time line information provided by the Chicago Fire Department and information provided from the NIOSH Fire Fighter Fatality Investigation and Prevention Program Report \cite{NIOSH:Bowyer}.

@@

\bigskip 
\begin{minipage}{.8\linewidth}
\centering
\captionof{table}{Fire Event Timeline}\label{tab:fire_info}
\begin{tabular}{c m{0.25in}}
\toprule[1.5pt]
Incident Time & Fire Behavior / Fireground Operation  \\
\midrule
17:15    & Reported working fire with heavy smoke coming from the front and rear of the structure's attic. \\
17:16    & Dispatch for a {\bf Still Alarm} \\
17:20    & Attic self ventilation - heavy smoke out of front attic window and heavy smoke and fire out of rear. \\  
17:21:00 & Engine 123 Capt. and FF on 2$^{nd}$ floor with 1 3/4-in hose line. \\
17:21:30 & 1$^{st}$ floor exterior porch door opened. \\
17:23    & Heavy fire report in rear stairwell and enclosed porch. No response from Engine 123: Capt. or FF. \\
17:24    & Water applied to attic from the rear exterior position. \\
17:27    & IC calls a Mayday \\
17:29    & Engine 123 Capt. carried out of structure. \\
\bottomrule[1.25pt]
\end{tabular}\par
\big-skip
\end{minipage}

@@

The development of the model geometry and development of the input file is straight forward with guidance from the FDS users' manual \cite{FDS_Users_Guide} or third party software.  Engineering judgment, graduate level knowledge of fire dynamics, and modeling experience are required to  translate the information collected from the scene into appropriate and realistic computational information.  Examples of these requirements include:
\begin{itemize}
\item the computational cell size
\item model domain and boundaries conditions
\item the heat release rate (HRR) of the ignition source/initial fire
\item physical and thermal properties of fuels such as interior finish and furnishings
	\begin{itemize}
	\item density
	\item specific heat
	\item thermal conductivity
	\item ignition temperature
	\item heat of combustion
	\item heat release rate
	\item mass loss rate under various fire conditions
	\item pyrolyzate/combustion product emissions
	\item response to thermal exposures
	\item activation energy
	\end{itemize}
\end{itemize}
Since there is no all-inclusive, universal fire modeling input data base this type of information must be gathered from literature or generated for a given incident. For the simulations performed for this study, Table ~\ref{tab:mod_param} includes the relevant model input parameters. The sections below describe these inputs.

\bigskip
\begin{minipage}{.8\linewidth}
\centering
\captionof{table}{Relevant Modeling Input Parameters}\label{tab:mod_param}
\begin{tabular}{lll}
\toprule[1.5pt]
Parameter & Description & Reference  \\
\midrule
 Simulation Time           & 5 minutes                             & -- \\[.25cm]
 Grid Size                 & 10 cm                                 & Section~\ref{mesh} \\ [.25cm]
 Ambient Temperature       & 25 $^{\circ}$C                        & -- \\ [.1cm]
\multirow{4}{*}{Reaction: Wood}  & Formula: $\C\H_{1.7}\O_{0.74}\N_{0.002}$              & \multirow{4}{*}{Section~\ref{fire}} \\
                           & CO Yield: 0.004 \; ${\mathrm{kg_{CO}}/\mathrm{kg_{wood}}}$  & \\
                           & Soot Yield: 0.015 \; ${\mathrm{kg_{C}}/\mathrm{kg_{wood}}}$ & \\
                           & Heat of Combustion: 16400 kJ/kg                             & \\ [.25cm]
\multirow{2}{*}{Peak HRR}  & Attic Fire: 9 MW                                            & \multirow{2}{*}{Section~\ref{fire}}\\ 
                           & Porch Fire: 3.3 MW                                          & \\ [.25cm]                     
\multirow{3}{*}{Material: Wood}    & k: 0.12 \; \si{W/(m.K)})        & \multirow{3}{*}{Section~\ref{matl}} \\
                                   & $\rho$: 510 \; kg/m$^3$         & \\
                                   & c$_{p}$: 1.38 \; \si{kJ/(kg.K)} & \\[.25cm]
\multirow{3}{*}{Material: Gypsum}  & k: 0.48 \; \si{W/(m.K)}         &  \multirow{3}{*}{Section~\ref{matl}}\\ 
                                   & $\rho$: 1440 \; kg/m$^3$        & \\ 
                                   & c$_{p}$: 0.84 \; \si{kJ/(kg.K)} & \\
\bottomrule[1.25pt]
\end{tabular}\par
\bigskip
\end{minipage}

\section{Geometry}
\label{geom}
The structure was a 2 $\tfrac{1}{2}$ story residential structure with balloon frame construction and a wood gable roofing system.  The structure had a 17.15 m (56 ft) by 6.15 m (20 ft) footprint. This results in a floor area of approximately 105 m$^2$ (1100 ft$^2$) per floor and a main living area of approximately 110 m$^2$ (2200 ft$^2$). Interior walls were constructed from wood frame dimensioned lumber with 1.27 cm (0.5 in) gypsum board over top. Additionally there was a half-story attic that was used to house the structure's hot water heater. There were three sets of stairs in the structure: one front stairwell that connected the first and second levels, one rear stairwell that connected the first and second levels and a second rear set of stairs which provided access to the attic. Fig.~(\ref{fig:alpha_ex}) shows the post-incident exterior of the alpha side of the structure. Note that the majority of the damage is to the structure's attic.

>>>>

\chapter{Introduction}
Part of the mission of the Fire Research Division at the National Institute of Standards and Technology (NIST) is to develop and apply technology, measurements and standards, improve the understanding of the behavior, prevention and control of fires to enhance fire fighting operations and equipment, fire suppression, fire investigations, and disaster response. NIST has previously used the Fire Dynamics Simulator (FDS) to provide insight on the fire development and thermal conditions of other fatality fires~\cite{Madrzykowski:1,Iowa,Texas,Cook_County,Grosshandler:Station,Bryner:Charleston,barowy:texas}. The overriding objective of all of these studies is to improve the safety of fire fighters.

@@

\chapter{Fire Incident Summary}
The following account of events is based on information provided by NIST physical measurements on-scene, NIST photographs of the scene, time line information provided by the Chicago Fire Department and information provided from the NIOSH Fire Fighter Fatality Investigation and Prevention Program Report~\cite{NIOSH:Bowyer}.

@@

\bigskip 
\begin{minipage}{.8\linewidth}
\centering
\captionof{table}{Fire Event Timeline}\label{tab:fire_info}
\begin{tabular}{c m{0.25in}}
\toprule[1.5pt]
Incident Time & Fire Behavior / Fireground Operation  \\
\midrule
17:15    & Reported working fire with heavy smoke coming from the front and rear of the structure's attic. \\
17:16    & Dispatch for a {\bf Still Alarm} \\
17:20    & Attic self ventilation - heavy smoke out of front attic window and heavy smoke and fire out of rear. \\  
17:21:00 & Engine 123 Capt. and FF on 2$^{nd}$ floor with 1 3/4-in hose line. \\
17:21:30 & 1$^{st}$ floor exterior porch door opened. \\
17:23    & Heavy fire report in rear stairwell and enclosed porch. No response from Engine 123: Capt. or FF. \\
17:24    & Water applied to attic from the rear exterior position. \\
17:27    & IC calls a Mayday \\
17:29    & Engine 123 Capt. carried out of structure. \\
\bottomrule[1.25pt]
\end{tabular}\par
\bigskip
\end{minipage}

@@

The development of the model geometry and development of the input file is straight forward with guidance from the FDS User's Guide  \cite{FDS_Users_Guide} or third party software.  Engineering judgment, graduate level knowledge of fire dynamics, and modeling experience are required to  translate the information collected from the scene into appropriate and realistic computational information.  Examples of these requirements include:
\begin{itemize}
\item the computational cell size
\item model domain and boundaries conditions
\item the heat release rate (HRR) of the ignition source/initial fire
\item physical and thermal properties of fuels such as interior finish and furnishings
\end{itemize}
Since there is no all-inclusive, universal fire modeling input data base this type of information must be gathered from literature or generated for a given incident. For the simulations performed for this study, Table ~\ref{tab:mod_param} includes the relevant model input parameters. The sections below describe these inputs.

\bigskip
\begin{minipage}{.8\linewidth}
\captionof{table}{Relevant Modeling Input Parameters}\label{tab:mod_param}
\begin{tabular}{lll}
\toprule[1.5pt]
Parameter & Description & Reference  \\
\midrule
 Simulation Time           & 5 minutes                             & -- \\[.25cm]
 Grid Size                 & 10 cm                                 & Section~\ref{mesh} \\ [.25cm]
 Ambient Temperature       & 25 $^{\circ}$C                        & -- \\ [.1cm]
\multirow{4}{*}{Reaction: Wood}  & Formula: $\C\H_{1.7}\O_{0.74}\N_{0.002}$              & \multirow{4}{*}{Section~\ref{fire}} \\
                           & CO Yield: 0.004 \; ${\mathrm{kg_{CO}}/\mathrm{kg_{wood}}}$  & \\
                           & Soot Yield: 0.015 \; ${\mathrm{kg_{C}}/\mathrm{kg_{wood}}}$ & \\
                           & Heat of Combustion: 16400 kJ/kg                             & \\ [.25cm]
\multirow{2}{*}{Peak HRR}  & Attic Fire: 9 MW                                            & \multirow{2}{*}{Section~\ref{fire}}\\ 
                           & Porch Fire: 3.3 MW                                          & \\ [.25cm]                     
\multirow{3}{*}{Material: Wood}    & k: 0.12 \; \si{W/(m.K)})        & \multirow{3}{*}{Section~\ref{matl}} \\
                                   & $\rho$: 510 \; kg/m$^3$         & \\
                                   & c$_{p}$: 1.38 \; \si{kJ/(kg.K)} & \\[.25cm]
\multirow{3}{*}{Material: Gypsum}  & k: 0.48 \; \si{W/(m.K)}         &  \multirow{3}{*}{Section~\ref{matl}}\\ 
                                   & $\rho$: 1440 \; kg/m$^3$        & \\ 
                                   & c$_{p}$: 0.84 \; \si{kJ/(kg.K)} & \\
\bottomrule[1.25pt]
\end{tabular}\par
\bigskip
\end{minipage}

\section{Geometry}
\label{geom}
The structure was a 2 1/2 story residential structure with balloon frame construction and a wood gable roofing system.  The structure had a 17.15 m (56 ft) by 6.15 m (20 ft) footprint. This results in a floor area of approximately 105 m$^2$ (1100 ft$^2$) per floor and a main living area of approximately 110 m$^2$ (2200 ft$^2$). Interior walls were constructed from wood frame dimensioned lumber with 1.27 cm (0.5 in) gypsum board over top. Additionally there was a half-story attic that was used to house the structure's hot water heater. There were three sets of stairs in the structure: one front stairwell that connected the first and second levels, one rear stairwell that connected the first and second levels and a second rear set of stairs which provided access to the attic. Fig.~(\ref{fig:alpha_ex}) shows the post-incident exterior of the alpha side of the structure. Note that the majority of the damage is to the structure's attic.
